A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-07-09 text/manuscript.Rmd
add second hallmarks reference

Other studies aim at providing a comprehensive view of mutations, gene
expression and genomic signatures, with the express goal of understanding the
common themes of all cancer independent of tissue of origin. Such a view of
cancer as a malady affecting common cellular processes has long been a goal of
the field and is perhaps best characterized in an essay by @weinberg2000 (see also
Discussion). The first of these studies, (the
“Pan-Cancer-12” analysis) identified 11 subtypes from 12 cancer types,
using integrative analysis with co-equal weighting of gene expression,
methylation, copy number and proteomics data
[@hoadleyMultiplatformAnalysis2014]. The principal finding was that
tissue-of-origin is the predominant driving factor, though a significant
proportion of tumors ($10\%$) could be reclassified independent of
tissue-of-origin. In a second study involving 33 cancer types and a much greater
number of tumors using a newer algorithm (iCluster)
[@shenIntegrativeClusteringMultiple2009], the authors identify 28 clusters that
could be further subdivided into organ specific groups, including
pan-gastrointestinal, pan-gynecological, pan-squamous, pan-gynecological/
squamous and pan-kidney [@hoadleyCellofOriginPatternsDominate2018].

>>>>

Other studies aim at providing a comprehensive view of mutations, gene
expression and genomic signatures, with the express goal of understanding the
common themes of all cancer independent of tissue of origin. Such a view of
cancer as a malady affecting common cellular processes has long been a goal of
the field and is perhaps best characterized in essays by @weinberg2000; @hallmarks_II (see also
Discussion). The first of these studies, (the
“Pan-Cancer-12” analysis) identified 11 subtypes from 12 cancer types,
using integrative analysis with co-equal weighting of gene expression,
methylation, copy number and proteomics data
[@hoadleyMultiplatformAnalysis2014]. The principal finding was that
tissue-of-origin is the predominant driving factor, though a significant
proportion of tumors ($10\%$) could be reclassified independent of
tissue-of-origin. In a second study involving 33 cancer types and a much greater
number of tumors using a newer algorithm (iCluster)
[@shenIntegrativeClusteringMultiple2009], the authors identify 28 clusters that
could be further subdivided into organ specific groups, including
pan-gastrointestinal, pan-gynecological, pan-squamous, pan-gynecological/
squamous and pan-kidney [@hoadleyCellofOriginPatternsDominate2018].
